---
title: "Life Expectancy across Countinents and Countries"
author: "Rajesh Ekkaladevi"
date: "3/29/2020"
output: 
    ioslides_presentation:
        smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Data
Exploring life expectancy across Continents and Countries based on GDP and Population as provided by the **gapminder** dataset.

```{r, echo=FALSE, results=FALSE, error=FALSE, warning=FALSE, message=FALSE}
library(plotly)
library(tidyverse)
library(gapminder)
library(RColorBrewer)
```

### Data Structure

```{r, echo=FALSE, results=TRUE, error=FALSE, warning=FALSE, message=FALSE}
gapminder %>% str()
```

***
### Sample Data (year: 2007)

```{r, echo=FALSE, results=TRUE, error=FALSE, warning=FALSE, message=FALSE}
gm <- gapminder %>% 
    filter(continent == 'Asia' & 
        year == 2007 &
        !is.na(lifeExp)
    ) %>%
    mutate(
        country = fct_reorder(country, gdpPercap),
        life_exp = round(lifeExp),
        gdp_lvl = cut(gdpPercap, breaks=quantile(.$gdpPercap, seq(0, 1, by=0.1)), labels=seq(1:10), include.lowest=TRUE),
        pop_lvl = cut(pop, breaks=quantile(.$pop, seq(0, 1, by=0.2)), labels=seq(1:5), include.lowest=TRUE)
    )
gm %>% head(10)
```

## Plot - Asia
**Life Expectancy by Country with GDP (dot color) and Population (dot size)**

```{r, echo=FALSE, results=TRUE, error=FALSE, warning=FALSE, message=FALSE}
gdp_lvl <- gm %>% arrange(gdp_lvl) %>% select(gdp_lvl) %>% unique()
pop_lvl <- gm %>% arrange(pop_lvl) %>% select(pop_lvl) %>% unique()
gdp_col <- c(colorRampPalette(c("red", "yellow", "green"))(10))

z <- ggplot(gm, aes(y=life_exp, x=country, color=gdp_lvl, size=pop_lvl)) + 
    geom_point(alpha=0.8, show.legend=FALSE) +
    scale_color_manual(labels=gdp_lvl, values=gdp_col) +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1),
        axis.ticks = element_blank(),
        panel.border = element_blank()
    ) +
    labs(x="Country", y="Life Expectancy", 
        colour="GDP Level", size="Population Level")
        
ggplotly(z)
```

## Plot - Africa
**Life Expectancy by Country with GDP (dot color) and Population (dot size)**

```{r, echo=FALSE, results=TRUE, error=FALSE, warning=FALSE, message=FALSE}
gm <- gapminder %>% 
    filter(continent == 'Africa' & 
        year == 2007 &
        !is.na(lifeExp)
    ) %>%
    mutate(
        country = fct_reorder(country, gdpPercap),
        life_exp = round(lifeExp),
        gdp_lvl = cut(gdpPercap, breaks=quantile(.$gdpPercap, seq(0, 1, by=0.1)), labels=seq(1:10), include.lowest=TRUE),
        pop_lvl = cut(pop, breaks=quantile(.$pop, seq(0, 1, by=0.2)), labels=seq(1:5), include.lowest=TRUE)
    )
    
gdp_lvl <- gm %>% arrange(gdp_lvl) %>% select(gdp_lvl) %>% unique()
pop_lvl <- gm %>% arrange(pop_lvl) %>% select(pop_lvl) %>% unique()
gdp_col <- c(colorRampPalette(c("red", "yellow", "green"))(10))

z <- ggplot(gm, aes(y=life_exp, x=country, color=gdp_lvl, size=pop_lvl)) + 
    geom_point(alpha=0.8, show.legend=FALSE) +
    scale_color_manual(labels=gdp_lvl, values=gdp_col) +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1),
        axis.ticks = element_blank(),
        panel.border = element_blank()
    ) +
    labs(x="Country", y="Life Expectancy", 
        colour="GDP Level", size="Population Level")
        
ggplotly(z)
```

## Plot - Americas
**Life Expectancy by Country with GDP (dot color) and Population (dot size)**

```{r, echo=FALSE, results=TRUE, error=FALSE, warning=FALSE, message=FALSE}
gm <- gapminder %>% 
    filter(continent == 'Americas' & 
        year == 2007 &
        !is.na(lifeExp)
    ) %>%
    mutate(
        country = fct_reorder(country, gdpPercap),
        life_exp = round(lifeExp),
        gdp_lvl = cut(gdpPercap, breaks=quantile(.$gdpPercap, seq(0, 1, by=0.1)), labels=seq(1:10), include.lowest=TRUE),
        pop_lvl = cut(pop, breaks=quantile(.$pop, seq(0, 1, by=0.2)), labels=seq(1:5), include.lowest=TRUE)
    )
    
gdp_lvl <- gm %>% arrange(gdp_lvl) %>% select(gdp_lvl) %>% unique()
pop_lvl <- gm %>% arrange(pop_lvl) %>% select(pop_lvl) %>% unique()
gdp_col <- c(colorRampPalette(c("red", "yellow", "green"))(10))

z <- ggplot(gm, aes(y=life_exp, x=country, color=gdp_lvl, size=pop_lvl)) + 
    geom_point(alpha=0.8, show.legend=FALSE) +
    scale_color_manual(labels=gdp_lvl, values=gdp_col) +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1),
        axis.ticks = element_blank(),
        panel.border = element_blank()
    ) +
    labs(x="Country", y="Life Expectancy", 
        colour="GDP Level", size="Population Level")
        
ggplotly(z)
```

## Plot - Europe
**Life Expectancy by Country with GDP (dot color) and Population (dot size)**

```{r, echo=FALSE, results=TRUE, error=FALSE, warning=FALSE, message=FALSE}
gm <- gapminder %>% 
    filter(continent == 'Europe' & 
        year == 2007 &
        !is.na(lifeExp)
    ) %>%
    mutate(
        country = fct_reorder(country, gdpPercap),
        life_exp = round(lifeExp),
        gdp_lvl = cut(gdpPercap, breaks=quantile(.$gdpPercap, seq(0, 1, by=0.1)), labels=seq(1:10), include.lowest=TRUE),
        pop_lvl = cut(pop, breaks=quantile(.$pop, seq(0, 1, by=0.2)), labels=seq(1:5), include.lowest=TRUE)
    )
    
gdp_lvl <- gm %>% arrange(gdp_lvl) %>% select(gdp_lvl) %>% unique()
pop_lvl <- gm %>% arrange(pop_lvl) %>% select(pop_lvl) %>% unique()
gdp_col <- c(colorRampPalette(c("red", "yellow", "green"))(10))

z <- ggplot(gm, aes(y=life_exp, x=country, color=gdp_lvl, size=pop_lvl)) + 
    geom_point(alpha=0.8, show.legend=FALSE) +
    scale_color_manual(labels=gdp_lvl, values=gdp_col) +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1),
        axis.ticks = element_blank(),
        panel.border = element_blank()
    ) +
    labs(x="Country", y="Life Expectancy", 
        colour="GDP Level", size="Population Level")
        
ggplotly(z)
```

## Plot - Oceania
**Life Expectancy by Country with GDP (dot color) and Population (dot size)**

```{r, echo=FALSE, results=TRUE, error=FALSE, warning=FALSE, message=FALSE}
gm <- gapminder %>% 
    filter(continent == 'Oceania' & 
        year == 2007 &
        !is.na(lifeExp)
    ) %>%
    mutate(
        country = fct_reorder(country, gdpPercap),
        life_exp = round(lifeExp),
        gdp_lvl = cut(gdpPercap, breaks=quantile(.$gdpPercap, seq(0, 1, by=0.1)), labels=seq(1:10), include.lowest=TRUE),
        pop_lvl = cut(pop, breaks=quantile(.$pop, seq(0, 1, by=0.2)), labels=seq(1:5), include.lowest=TRUE)
    )
    
gdp_lvl <- gm %>% arrange(gdp_lvl) %>% select(gdp_lvl) %>% unique()
pop_lvl <- gm %>% arrange(pop_lvl) %>% select(pop_lvl) %>% unique()
gdp_col <- c(colorRampPalette(c("lightgreen", "green"))(2))

z <- ggplot(gm, aes(y=life_exp, x=country, color=gdp_lvl, size=pop_lvl)) + 
    geom_point(alpha=0.8, show.legend=FALSE) +
    scale_color_manual(labels=gdp_lvl, values=gdp_col) +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1),
        axis.ticks = element_blank(),
        panel.border = element_blank()
    ) +
    labs(x="Country", y="Life Expectancy", 
        colour="GDP Level", size="Population Level")
        
ggplotly(z)
```
