---
title: "Developing Data Products - Programming Assignment - Week 2"
author: "Rajesh Ekkaladevi"
date: "Mar 27, 2020"
output:
  html_document:
    highlight: espresso
    theme: journal
  pdf_document:
    highlight: espresso
always_allow_html: true
---

## Executive Summary

The goal of the exercise is to create a map using Leaflet.

A map describing earth quakes info across the world between years 2000 and 2020 is created using Leaflet.  The map provides a hover feature that would display the magnitude, year and also place of the earth quake.  When clicked on the circle a standing out label with the info is diplayed.  The magnitude of the earth quakes can be selected using checkbox on the top of the map.

The earth quake data is obtained from https://earthquake.usgs.gov website and is downloaded ahead into the project directory.

## Map
```{r, echo=FALSE, results=FALSE, error=FALSE, warning=FALSE, message=FALSE}
library(leaflet)
library(tidyverse)
library(lubridate)
library(RColorBrewer)
```

```{r, echo=FALSE, results=TRUE, error=FALSE, warning=FALSE, message=FALSE}
# full data and color palette.
data <- read_csv("world_earthquakes.csv") %>% 
    filter(mag >= 5) %>% 
    mutate(mag_lvl = cut(mag, breaks=seq(5,10), labels=paste0("mag", 5:9), right=FALSE))
    
mag5 <- data %>% filter(mag_lvl == 'mag5')
mag6 <- data %>% filter(mag_lvl == 'mag6')
mag7 <- data %>% filter(mag_lvl == 'mag7')
mag8 <- data %>% filter(mag_lvl == 'mag8')
mag9 <- data %>% filter(mag_lvl == 'mag9')
    
colorpal <- colorNumeric("YlOrRd", data$mag)

# draw map, center view to north america and plot layered circles with magnitude info.
leaflet() %>% 
    setView(lng = -99, lat = 42, zoom = 3)  %>%
    addTiles() %>% 
    addCircles(data = mag5,
        lat = ~ latitude, 
        lng = ~ longitude, 
        color = "#777777",
        fillColor = ~colorpal(mag), 
        fillOpacity = 0.5,
        radius = ~ exp(sqrt(mag))*mag*1000, 
        weight = 1, 
        group = "5",
        popup = ~as.character(sprintf("Magnitude: %s<br>Date: %s<br>Place: %s", mag, year(time), place)), 
        label = ~as.character(sprintf("Magnitude: %s, Date: %s, Place: %s", mag, year(time), place))
    ) %>% 
    addCircles(data = mag6,
        lat = ~ latitude, 
        lng = ~ longitude, 
        color = "#777777",
        fillColor = ~colorpal(mag), 
        fillOpacity = 0.5,
        radius = ~ exp(sqrt(mag))*mag*1000, 
        weight = 1, 
        group = "6",
        popup = ~as.character(sprintf("Magnitude: %s<br>Date: %s<br>Place: %s", mag, year(time), place)), 
        label = ~as.character(sprintf("Magnitude: %s, Date: %s, Place: %s", mag, year(time), place))
     ) %>% 
    addCircles(data = mag7,
        lat = ~ latitude, 
        lng = ~ longitude, 
        color = "#777777",
        fillColor = ~colorpal(mag), 
        fillOpacity = 0.5,
        radius = ~ exp(sqrt(mag))*mag*1000, 
        weight = 1, 
        group = "7",
        popup = ~as.character(sprintf("Magnitude: %s<br>Date: %s<br>Place: %s", mag, year(time), place)), 
        label = ~as.character(sprintf("Magnitude: %s, Date: %s, Place: %s", mag, year(time), place))
     ) %>% 
    addCircles(data = mag8,
        lat = ~ latitude, 
        lng = ~ longitude, 
        color = "#777777",
        fillColor = ~colorpal(mag), 
        fillOpacity = 0.5,
        radius = ~ exp(sqrt(mag))*mag*1000, 
        weight = 1, 
        group = "8",
        popup = ~as.character(sprintf("Magnitude: %s<br>Date: %s<br>Place: %s", mag, year(time), place)), 
        label = ~as.character(sprintf("Magnitude: %s, Date: %s, Place: %s", mag, year(time), place))
     ) %>% 
    addCircles(data = mag9,
        lat = ~ latitude, 
        lng = ~ longitude, 
        color = "#777777",
        fillColor = ~colorpal(mag), 
        fillOpacity = 0.5,
        radius = ~ exp(sqrt(mag))*mag*1000, 
        weight = 1, 
        group = "9",
        popup = ~as.character(sprintf("Magnitude: %s<br>Date: %s<br>Place: %s", mag, year(time), place)), 
        label = ~as.character(sprintf("Magnitude: %s, Date: %s, Place: %s", mag, year(time), place))
    ) %>% 
    addLayersControl(overlayGroups = c("5", "6", "7", "8", "9"), options = layersControlOptions(collapsed = FALSE)) %>% 
    addLegend(pal = colorpal, values = data$mag, position = "bottomleft")
```
